\LARGE \bf Eye tracking with Educational Robots: A Cautionary Tale

### 2014-09-29 paper.tex
Added biblio. 2 entries are missing

On the other hand, educational robots can be of great use in making
young school kids interested in technological developments
~\cite{2, 17}. It is important to know about the key socio-cognitive
features of this subset of human-robot interaction because these
features let the researchers design better interactive scenarios in the
future. These features include the conceptual mapping of the sensor data
visualization and the robot's function and the generalizing the concepts
to solve a given problem in a contextual learning environment. ~

Thymio II~\cite{10} can provide salient effects onto our visual attention
through the LEDs. We can use this feature to guide the users' attention
in a way that the effectiveness of the interaction is maximized. In the
case of interaction with the educational robots the effect of the
interaction is to make the users understand the technical aspects of the
robot and to make the users be able to transfer it to different domains.

We present a study with combining eye tracking and human robot
interaction (HRI) in an educational setting. ~Eye tracking data is of
high temporal precision as compared to some of the physiological data
sources such as fMRI or other biometric devices used to measure the
palpitation and the body temperature. The eye tracking devices are
portable enough as opposed to the bulky apparatus used in fMRI~\cite{9}.
Eye tracking data provides more direct access to the users' attention
than other portable and physiological data sources such as EEG; this
makes eye-tracking data a pertinent source to measure attention.

@@

Our working hypothesis is that with careful mapping between the LED
visualizations and the behavior of the robot we can induce the
understanding of the functionality of the robot very easily. We propose
to use the gaze data to verify this hypothesis. With eye-tracking data
we can judge whether the mapping between the behavior and visual
actuators is useful or not for the participants' understanding.
Moreover, we hypothesize that the good performers will be able to
anticipate the robot's motion. Eye-tracking data can also be useful to
verify the anticipation hypothesis. ~

@@

Several scholars related gaze patterns with level of expertise. \cite{4}
in an air traffic-monitoring task found out that the experts looked less
at the scenario-specific information than novices.
\cite{3, 7, 16} studied the effect of expertise on the gaze
patterns in different surgical tasks and concluded that experts look
less at the instruments than the novices, instead they focus more on the
task specific areas. \cite{11} showed that expert chess players pay more
attention to the relative positions of the pieces, rather than the
individual pieces, than novice chess players. \cite{1} also studied the
difference between experts and novice chess players in a checkmate
avoidance task and concluded that the experts have more gaze falling on
the important squares than the novices. In a program-debugging task, \cite{12} showed that the expert programmers scan through all the lines
in the program faster than the novices. In a collaborative Tetris game,
\cite{5} showed that experts pay more attention to the stack of
Tetronimoes while novices allocate more attention to the new pieces
falling from the top.

Our previous work also show a clear relation between gaze patterns and
task-based performance. In a pair-programming task, \cite{6} showed that
the good performing pairs have more synchronized gaze on different parts
of a program than the bad performing pairs. In a similar task, \cite{13}
showed that the good performing pairs pay more attention to the
data-flow of the program than the poor performing pairs. Moreover,
\cite{14} showed that while describing the functionality of a program the
well performing teams had more gaze on the variable modification parts
in the program while poor performing teams have equal distribution of
gaze on different parts of the program during similar phase of the task.
Thus we see that previous research provides insights about the
relationship between the gaze patterns and the behavioral and task-based
performance indicators in diverse scenarios.

It must be noted that to our best knowledge only very few studies have
investigated how eye-tracking could be used to assess HRI. In an
eye-tracking study, in the context of human robot interaction, \cite{15}
found that when the robot provides a matching visual and verbal
reference to an object the participants were more likely to look at the
correct objects. This result was consistent irrespectively of the
ambiguity of the situation.

@@

Thymio II is a small robot designed for education for 6 to 16 year old
children~\cite{8, 10}. Users can interact with it via buttons and
distance sensors. The robot can display its states via many LEDs and its
motion. We programmed it to show the value of the front IR-sensor on its
top. The display consists of 8 LEDs (Figure~\ref{thymio}).

~In the experiment, there were 3 sensor data visualization conditions.
~For the TRUE visualization condition, the number of illuminated LEDs,
on the top of the robot, is proportional to the intensity measured by
the IR-sensor. When the intensity of the reflected IR crosses a
predefined threshold, the robot does a 90° turn to its right. For the
two other conditions, the behavior is the same, simply the display
changes. In the RANDOM condition, the display shows as if the robot was
detecting the obstacle at random moments, while in the NONE
visualization condition nothing is shown on top of the robot. The robot
has a fitting for a pencil that we used for the drawing of its
trajectory.

@@

The most common form of eye-tracking data aggregation are the fixations.
Fixations depict the periods in the user's observation when the user
attends a relatively small part of the visual stimulus for a relatively
longer period of time. The various fixation detection algorithms are
summarized in~\cite{18}. Following are the two gaze measures, derived
from the fixations, which we used in the present work.

@@

We demonstrated that eye-tracking data can provide several process
variables to assess not only the task-based performance of the
participants but also to evaluate HRI. The prime advantage of using
mobile eye-tracking is that the users' interaction with the robot can
still be ecologically valid. On the other hand, using fMRI for
evaluating HRI reduces the interaction to a video of the robot only
~\cite{9}.

@@

Another plausible explanation for the participants in NONE and RANDOM condition
being better performers than the TRUE condition during the interaction phase
comes from the transparency of the robot's behavior~\cite{20}. The robot's
behavior was not at all transparent in the former two conditions. This might
have caused the participants to put more effort in understanding the behavior.
Therefore, they performed better.

>>>>

On the other hand, educational robots can be of great use in making young school
kids interested in technological developments~\cite{cooper1999robots,
wyffels2010building}. It is important to know about the key socio-cognitive
features of this subset of human-robot interaction because these features let
the researchers design better interactive scenarios in the future. These
features include the conceptual mapping of the sensor data visualization and the
robot's function and the generalizing the concepts to solve a given problem in a
contextual learning environment. ~

Thymio II~\cite{riedo2012two} can provide salient effects onto our visual
attention through the LEDs. We can use this feature to guide the users'
attention in a way that the effectiveness of the interaction is maximized. In
the case of interaction with the educational robots the effect of the
interaction is to make the users understand the technical aspects of the robot
and to make the users be able to transfer it to different domains.

We present a study with combining eye tracking and human robot interaction (HRI)
in an educational setting. ~Eye tracking data is of high temporal precision as
compared to some of the physiological data sources such as fMRI or other
biometric devices used to measure the palpitation and the body temperature. The
eye tracking devices are portable enough as opposed to the bulky apparatus used
in fMRI~\cite{rosenthal2013neural}.  Eye tracking data provides more direct
access to the users' attention than other portable and physiological data
sources such as EEG; this makes eye-tracking data a pertinent source to measure
attention.

@@

Our working hypothesis is that with careful mapping between the LED
visualizations and the behavior of the robot we can induce the understanding of
the functionality of the robot very easily. We propose to use the gaze data to
verify this hypothesis. With eye-tracking data we can judge whether the mapping
between the behavior and visual actuators is useful or not for the participants'
understanding.  Moreover, we hypothesize that the good performers will be able
to anticipate the robot's motion. Eye-tracking data can also be useful to verify
the anticipation hypothesis.

@@

Several scholars related gaze patterns with level of expertise.
\cite{hasse2012measure} in an air traffic-monitoring task found out that the
experts looked less at the scenario-specific information than novices.
\cite{eivazi2012gaze, law2004eye, tien2010measuring} studied the effect of
expertise on the gaze patterns in different surgical tasks and concluded that
experts look less at the instruments than the novices, instead they focus more
on the task specific areas. \cite{reingold2001visual} showed that expert chess
players pay more attention to the relative positions of the pieces, rather than
the individual pieces, than novice chess players. \cite{blignaut2008visual} also
studied the difference between experts and novice chess players in a checkmate
avoidance task and concluded that the experts have more gaze falling on the
important squares than the novices. In a program-debugging task,
\cite{sharif2012eye} showed that the expert programmers scan through all the
lines in the program faster than the novices. In a collaborative Tetris game,
\cite{jermann2010using} showed that experts pay more attention to the stack of
Tetronimoes while novices allocate more attention to the new pieces falling from
the top.

Our previous work also show a clear relation between gaze patterns and
task-based performance. In a pair-programming task, \cite{jermann2012effects}
showed that the good performing pairs have more synchronized gaze on different
parts of a program than the bad performing pairs. In a similar task,
\cite{sharma2012gaze} showed that the good performing pairs pay more attention
to the data-flow of the program than the poor performing pairs. Moreover,
\cite{sharma2013understanding} showed that while describing the functionality of
a program the well performing teams had more gaze on the variable modification
parts in the program while poor performing teams have equal distribution of gaze
on different parts of the program during similar phase of the task.  Thus we see
that previous research provides insights about the relationship between the gaze
patterns and the behavioral and task-based performance indicators in diverse
scenarios.

It must be noted that to our best knowledge only very few studies have
investigated how eye-tracking could be used to assess HRI. In an eye-tracking
study, in the context of human robot interaction, \cite{staudte2009visual} found
that when the robot provides a matching visual and verbal reference to an object
the participants were more likely to look at the correct objects. This result
was consistent irrespectively of the ambiguity of the situation.

@@

Thymio II is a small robot designed for education for 6 to 16 year old
children~\cite{magnenat2012programming, riedo2012two}. Users can interact with
it via buttons and distance sensors. The robot can display its states via many
LEDs and its motion. We programmed it to show the value of the front IR-sensor
on its top. The display consists of 8 LEDs (Figure~\ref{thymio}).

In the experiment, there were 3 sensor data visualization conditions.  For the
TRUE visualization condition, the number of illuminated LEDs, on the top of the
robot, is proportional to the intensity measured by the IR-sensor. When the
intensity of the reflected IR crosses a predefined threshold, the robot does a
90° turn to its right. For the two other conditions, the behavior is the same,
simply the display changes. In the RANDOM condition, the display shows as if the
robot was detecting the obstacle at random moments, while in the NONE
visualization condition nothing is shown on top of the robot. The robot has a
fitting for a pencil that we used for the drawing of its trajectory.

@@

The most common form of eye-tracking data aggregation are the fixations.
Fixations depict the periods in the user's observation when the user
attends a relatively small part of the visual stimulus for a relatively
longer period of time. The various fixation detection algorithms are
summarized in~\cite{duchowski2007eye}. Following are the two gaze measures, derived
from the fixations, which we used in the present work.

@@

We demonstrated that eye-tracking data can provide several process
variables to assess not only the task-based performance of the
participants but also to evaluate HRI. The prime advantage of using
mobile eye-tracking is that the users' interaction with the robot can
still be ecologically valid. On the other hand, using fMRI for
evaluating HRI reduces the interaction to a video of the robot only
~\cite{rosenthal2013neural}.

@@

Another plausible explanation for the participants in NONE and RANDOM condition
being better performers than the TRUE condition during the interaction phase
comes from the transparency of the robot's behavior~\cite{??}. The robot's
behavior was not at all transparent in the former two conditions. This might
have caused the participants to put more effort in understanding the behavior.
Therefore, they performed better.
